Application of Micro-Rotors to Induce Crystal Nucleation: Challenges, Takeaways, and Future Prospects\\ PhD Thesis

### 2024-09-23 Latex_files/Thesis/chapters/chapter 5/main.tex
updates

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Calculation of error}
\label{sec:divergence}
To evaluate the above estimation of dimer orientation from scattering signal, we use a Brownian simulation of a dimer in the optical trap (Section~\ref{sec:brownian}) to compare estimated most probable reference orientation, derived from the dimer's scattering through Eq.~\eqref{eq:bayes}, with the dimer's known \emph{actual} orientation $\hat{\bf s}$. MSTM provides calculated light scattering from the simulated dimer $I(\hat{\bf s}, \theta)$ and we use \eqref{eq:scale} to obtain normalized values at each measurement angle $\theta_k$,  $y_1(\hat{\bf s})$, $y_2(\hat{\bf s})$, $y_3(\hat{\bf s})$, from which we obtain $p(\hat{\bf n}_{\alpha}\parallel y_1, y_2, y_3)$. Because we know the actual orientation $\hat{s}$ we can measure the error in the model's estimate by comparing the reference orientation closest to $\hat{s}$, denoted as $\hat{\bf n}_{best}$, with the most probable predicted orientation from Eq.~\eqref{eq:bayes}. An ideal result would be one where the probability distribution is 0 for every $\hat{\bf n}$ apart from $\hat{\bf n}_{best}$:
\begin{align}
	\label{eq:best}
	p_{best} = 
	\begin{cases}
		1 & \text{when $\hat{\bf n}_\alpha$ = $\hat{\bf n}_{best}$}\\
		0 & \text{anywhere else}
	\end{cases}
\end{align}
In reality the distribution from Eq.~\eqref{eq:bayes} will assign some non-zero
probability to every reference orientation, leading to some level 'confidence' in orientation prediction, which can be quantified by calculating the Kullback-Leibler divergence $K_l$ between the two distributions:
\begin{align}
	K_{l, \#}(p_{best}\parallel p(\hat{\bf n}_\alpha| y_1, y_2, y_3))
	= p_{best}\ln \left[\frac{p_{best}}{p(\hat{\bf n}_{best}| y_1,y_2,y_3))}
	\right]
	\label{eq;kullback}
\end{align}
where a larger value of $K_l$ indicates that our model is less confident in its
prediction of the dimer's orientation. The divergence $K_l$ thus illustrates the 'spread'
in the estimated dimer orientation probability --- a distribution strongly peaked at 
some value would give us more confidence in that value than a near-uniform distribution 
where the scattering measurement could imply a wide range of possible orientations --- 
but it does not directly indicate our estimates actual accuracy, that can be simply defined as the percentage of our estimations that are correct.

@@

We use the Brownian OT package developed by Fung~\textit{et~al} \cite{Vigilante2020} to simulate the motion of an asymmetric dimer (Figure~\ref{fig:dimer}a) within an optical trap. Brownian OT combines MSTM \cite{Mackowski1996} and ``Optical Tweezer Toolbox'' (\textit{ott}) \cite{Lenton2020} to simulate the motion of arbitrary
shaped sphere clusters. We simulate the motion of a dimer trapped in a highly focused Gaussian beam by calculating the optical forces imparted by the laser, and the Brownian force due to the surrounding fluid. MSTM provides the necessary T-matrix to compute the optical force via \textit{ott}. The Brownian force is found by computing the dimer's diffusion tensor according to the analytical solutions provided by Nir and Acrivos \cite{nir_acrivos_1973}. We simulated a polystyrene dimer ($n = 1.59$) in a suspension of water ($n_{med} = 1.33$) over the course of $1 \ s$ with a simulation time step of $1 \times 10^{-5} \ s$. We placed the dimer 4 microns below the trap focus at an angle $30^{\circ}$ from the horizontal, the resulting trajectory is shown below in Sec~\ref{sec:motion}. We chose these initial parameters because it demonstrates our model's performance in non steady state conditions.

@@

In the simulations of Vigilante~\emph{et~al}.\ \cite{Vigilante2020Brownian_OT}, trapped symmetrical dimers were investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment. However as seen in Figure~\ref{fig:motion} asymmetric dimers demonstrate dynamics that do not immediately achieve steady state. We chose to use asymmetric dimers as our benchmark due to this fact, as its orientational motion is far more complex than a symmetric dimer. In the future we hope to further investigate the motion of asymmetric dimers. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Accounting for sources of error in light scattering measurements}
When it comes to analysing light scattering from any size particle, error analysis becomes a significant factor. Typically this can be accounted for by averaging over long periods of time to get an assessment of the steady state conditions of the target particle. However in our case where we wish to know the instantaneous orientation, we instead have to rely on our understanding of how uncertainty can effect our model's performance. We identified two areas which are likely sources of error in our estimation: firstly, an incorrect modelling of the target particle, and secondly, signal noise arising from experimental factors. We highlight how we address these areas below. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subparagraph{Impact of incorrect dimer sizing}
\label{sec:lam}

@@

As can be seen from Fig~\ref{fig:size}  even the slightest change in size ratio makes a very significant difference to the performance of our model. This amounts to just over $100 \ nm$ in the dimer's overall size, yet results in our model being correct from over 90 \% of the time to now as low as 30 \%. This highlights the importance of correctly sizing trapped entities before performing any in depth analysis of the scattering pattern, as even the slightest deviation can have a serious impact. We addressed this by increasing the number of available reference orientations from 30 to 126 (following the same procedure as given by \cite{Rey2006} to evenly space out the coordinates) and increasing the weighting factor in Eq~\ref{eq:boltz}. While this didn't have a significant improvement on the overall accuracy of the model, in the worst case having a slight increase from 30.5 \% to 40.3 \%, it did help to significantly reduce the magnitude between our model's estimations and the dimer's motion as seen below in Fig~\ref{fig:refs}.
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.31\textwidth}
		\subcaption{}
		\includegraphics[width =\textwidth]{fig9a.png}
	\end{subfigure}
	\begin{subfigure}{0.31\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig9b.png}
	\end{subfigure}
	\begin{subfigure}{0.31\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig9c.png}
	\end{subfigure}
	\caption{Model estimates of orientation when neural net has been trained on dimer of size ratio: (a) 1:2 $[F(K_l)=11.756]$, (b) 1:2.05 $[F(K_l)=1.233]$, (c) 1:1.95 $[F(K_l)=2.128]$, ($n_{refs} = 126$)}
	\label{fig:refs}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Calculation of error}
\label{sec:divergence}
Using \eqref{eq:bayes} we get a snapshot estimation of the instantaneous
orientation of the trapped dimer based on a single measurement from a 
collection of detector signals. Of course in reality these detectors 
will be returning a discrete time series representative of the dimer's
trajectory within the optical trap.

Using the simulation code from \cite{Vigilante2020} we can replicate 
this trajectory and furthermore use \textit{mstm}to calculate the far field 
intensity produced by said dimer. At each time step we measure the signal
produced at each measurement angle and use our neural network to calculate
$p(y_k(\hat{\bf s})|\hat{\bf n}_\alpha, t)$.

Evaluating how well Eq.~\eqref{eq:bayes} performs in predicting $\hat{\bf n}$
is difficult because we have discretised the orientation space meaning 
there will always be some degree of inaccuracy in our prediction. The Kullback-Liebler divergence method allows us to compare the divergence 
between two probability, namely Eq.~\eqref{eq:bayes} and a case scenario.
\begin{align}
	D_{KL}(P\parallel p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}), t)) = \sum P(t) \frac{P(t)}{p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}), t)} 
\end{align}

Where $P(t)$ is the case scenario we are comparing against. In an ideal
case our model would correctly predict the dimer's orientation with 
100\% accuracy:
\begin{align}
	\label{eq:best}
	p_{best} = P(t) =
	\begin{cases}
		1 & \text{when $\hat{\bf n}_\alpha$ = $\hat{\bf n}_{best}$}\\
		0 & \text{when $\hat{\bf n}_\alpha$ $\neq$ $\hat{\bf n}_{best}$}
	\end{cases}
\end{align}

In reality the distribution from Eq.~\eqref{eq:bayes} will assign some 
non-zero probability to every reference orientation. Therefore the 
divergence of Eq.~\eqref{eq:bayes} at each time step is given as:
\begin{align}
	D_{KL}(p_{best}\parallel p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}), t))
	= \ln \left[\frac{1}{p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}), t)}
	\right]
	\label{eq;kullback}
\end{align}

where a larger value of $D_{KL}$ indicates that our model is less confident 
in its prediction of the dimer's orientation. The divergence $D_{KL}$ thus 
illustrates the 'spread' in the estimated dimer orientation probability 
-- a distribution strongly peaked at some value would give us more 
confidence in that value than a near-uniform distribution where the 
scattering measurement could imply a wide range of possible orientations 
-- but it does not directly indicate our estimates actual accuracy. 
In order to reflect the accuracy of each prediction by using a weighting
function such as:
\begin{align}
	\omega_\alpha = \frac{1}{2} (\hat{\bf n}_{best} \cdot \hat{\bf n}_\alpha+1)
\end{align}

Which is just the normalised dot product between our estimate and the
best possible estimate, so for highly inaccurate estimations $D_{KL}$
is far greater than if we narrowly missed the $\hat{\bf n}_{best}$. With
this in mind the divergence for a single measurement is:

\begin{align}
	D_{KL}(p_{best}\parallel p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}), t))
	= \ln \left[\frac{1}{p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}), t)\omega_\alpha}
	\right]
\end{align}

Later we will show how we set a baseline for evaluating the performance 
of \eqref{eq:bayes} across an entire trajectory and how it handles common
errors in a static light scattering setup.

@@

We use the Brownian OT package developed by Fung~\textit{et~al} 
\cite{Vigilante2020} to simulate the motion of an asymmetric dimer (Figure~\ref{fig:dimer}a) within an optical trap. We simulate the 
motion of a dimer trapped in a highly focused Gaussian beam by 
calculating the optical forces imparted by the laser, and the 
Brownian force due to the surrounding fluid. MSTM provides the 
necessary T-matrix to compute the optical force via \textit{ott}. 
The Brownian force is found by computing the dimer's diffusion 
tensor according to the analytical solutions provided by Nir and 
Acrivos \cite{Nir1973}. We simulated a polystyrene dimer ($n = 
1.59$) in a suspension of water ($n_{med} = 1.33$) over the 
course of $1 \ s$ with a simulation time step of $1 \times 10^{-5} 
\ s$. We placed the dimer 4 microns below the trap focus at an 
angle $30^{\circ}$ from the horizontal, the resulting trajectory 
is shown below in Sec~\ref{sec:motion}. We chose these initial 
parameters because it demonstrates our model's performance in non 
steady state conditions.

@@

In the simulations of Vigilante~\emph{et~al}.\ \cite{Vigilante2020}, trapped symmetrical dimers were investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment. However as seen in Figure~\ref{fig:motion} asymmetric dimers demonstrate dynamics that do not immediately achieve steady state. We chose to use asymmetric dimers as our benchmark due to this fact, as its orientational motion is far more complex than a symmetric dimer. In the future we hope to further investigate the motion of asymmetric dimers. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Accounting for sources of error in light scattering measurements}
When it comes to analysing light scattering from any size particle, error analysis becomes a significant factor. Typically this can be accounted for by averaging over long periods of time to get an assessment of the steady state conditions of the target particle. However in our case where we wish to know the instantaneous orientation, we instead have to rely on our understanding of how uncertainty can effect our model's performance. We identified two areas which are likely sources of error in our estimation: firstly, an incorrect modelling of the target particle, and secondly, signal noise arising from experimental factors. We highlight how we address these areas below. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subparagraph{Impact of incorrect dimer sizing}
\label{sec:lam}

@@

As can be seen from Fig~\ref{fig:size}  even the slightest change in size ratio makes a very significant difference to the performance of our model. This amounts to just over $100 \ nm$ in the dimer's overall size, yet results in our model being correct from over 90 \% of the time to now as low as 30 \%. This highlights the importance of correctly sizing trapped entities before performing any in depth analysis of the scattering pattern, as even the slightest deviation can have a serious impact. We addressed this by increasing the number of available reference orientations from 30 to 126 (following the same procedure as given by \cite{Reyuthor2006} to evenly space out the coordinates) and increasing the weighting factor in Eq~\ref{eq:boltz}. While this didn't have a significant improvement on the overall accuracy of the model, in the worst case having a slight increase from 30.5 \% to 40.3 \%, it did help to significantly reduce the magnitude between our model's estimations and the dimer's motion as seen below in Fig~\ref{fig:refs}.
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.31\textwidth}
		\subcaption{}
		\includegraphics[width =\textwidth]{fig9a.png}
	\end{subfigure}
	\begin{subfigure}{0.31\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig9b.png}
	\end{subfigure}
	\begin{subfigure}{0.31\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig9c.png}
	\end{subfigure}
	\caption{Model estimates of orientation when neural net has been trained on dimer of size ratio: (a) 1:2 $[F(K_l)=11.756]$, (b) 1:2.05 $[F(K_l)=1.233]$, (c) 1:1.95 $[F(K_l)=2.128]$, ($n_{refs} = 126$)}
	\label{fig:refs}
\end{figure}
